Approximation as Differentiation

### 2025-06-09 main/first-order.tex
Tweaks.

One might also like to allow the programmer to embed additional ``approximation points'' into composite
structures, by providing them with a \emph{lifting} monad $\synVar{L}$. Rather than the usual definition of
the type of lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could then define
(within the language) a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat}
\times \alpha)$, along with an approximating version of multiplication $*_{\text{appr}}:
\synVar{L}\,\synVar{nat} \times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$ expressible as
\[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\].

\noindent This would allow the programmer to emulate the short-circuiting behaviour of $\mathrm{strictOr}$
from \exref{strict-short-circuit}, or the short-circuiting multiplication used in \citet{perera22} for slicing
matrix convolutions, without baking this specific approximating behaviour into the interpretation of
individual operators.

@@

We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
vector space $\RR^n$ as paired with the constant family of tangent spaces at $\RR^n$ (also written $\RR^n$,
for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$.
Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
\partial f)$ corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp \partial
f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.

>>>>

One might also like to allow the programmer to embed additional ``approximation points'' into composite
structures, by providing them with a \emph{lifting} monad $\synVar{L}$. Rather than the usual definition of
the type of lists of natural numbers as $\mu\alpha.1 + \synVar{nat} \times \alpha$, one could (within the
language) define a type of ``approximation lists'' $\mu\alpha.\synVar{L}\,(1 + \synVar{L}\,\synVar{nat} \times
\alpha)$, along with an approximating version of multiplication $*_{\text{appr}}: \synVar{L}\,\synVar{nat}
\times \synVar{L}\,\synVar{nat} \to \synVar{L}\,\synVar{nat}$ expressible as
\[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\].

\noindent This would allow the programmer to emulate the short-circuiting behaviour of $\mathrm{strictOr}$
from \exref{strict-short-circuit}, or the short-circuiting multiplication used in \citet{perera22} for slicing
matrix convolutions, without baking this specific choice of approximation into the interpretation of
individual operators.

@@

We can see these ``optimised'' forward-mode maps as living in $\Fam(\FinVect)$ if we regard every finite
vector space $\RR^n$ as paired with the constant family of tangent spaces at $\RR^n$ (also written $\RR^n$,
for convenience), and each map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$ as the pair $(f_1, f_2)$.
Then the composition $(g, \partial g) \comp (f, \partial f) = (g \comp f, \reindex{\partial g}{f} \comp
\partial f)$ again corresponds exactly to the forward chain rule, since $(\reindex{\partial g}{f} \comp
\partial f)_x = {\partial g}_{f(x)} \comp {\partial f}_x$.
